---
title: "Base Operations Crime Data"
author: "Sarah Raza (sr6bf)"
date: '2022-12-02'
output: html_document
---

```{r, warning=FALSE,message=FALSE}
library(dplyr)
library(tidyverse)
library(reprex)
library(ggplot2)
library(RColorBrewer)
bcrime<-read.csv("/Users/sarahraza/Downloads/Boston_Crime_Data.csv")
```
Data cleaning
```{r}
bcrime = bcrime%>%select(-INCIDENT_NUMBER,-OFFENSE_CODE)
bcrime = bcrime%>%mutate(across(starts_with("Location"),~gsub("\\,", "", .)))
bcrime = bcrime%>%mutate(across(starts_with("Location"),~gsub("\\.", "", .)))
bcrime = bcrime%>%mutate(across(starts_with("Location"),~gsub("\\(", "", .)))
bcrime = bcrime%>%mutate(across(starts_with("Location"),~gsub("\\)", "", .)))
bcrime = na.omit(bcrime)
```
Data cleaning
```{r}
bcrime2 = bcrime%>%mutate(HAD_SHOOTING=ifelse(SHOOTING>0,"Y","N"))%>%select(-SHOOTING)
bcrime3 = bcrime2 %>% mutate(Crimes = OFFENSE_CODE_GROUP)
bcrime3$Crimes =  gsub('Simple Assault', 'Assault',
                  gsub('Aggravated Assault', 'Assault',
                  gsub('Property Lost', 'Property',
                  gsub('Aircraft','Vehicle',
                  gsub('Arson','Property',
                  gsub('Assembly or Gathering Violation','Property',
                  gsub('Auto Theft','Vehicle',
                  gsub('Auto Theft Recovery','Vehicle',
                  gsub('Biological Threat','Other',
                  gsub('Commercial Burglary','Theft',
                  gsub('Criminal Harassment','Harassment',
                  gsub('Disorderly Conduct','Harassment',
                  gsub('Embezzlement','Theft',
                  gsub('Confidence Games','Theft',
                  gsub('Fire Related Reports','Property',
                  gsub('Firearm Discovery','Weapon',
                  gsub('Fraud','Theft',
                  gsub('Police Theft Incidents','Theft',
                  gsub('Prisoner Related Incidents','Other',
                  gsub('Gambling','Theft',
                  gsub('Harbor Related Incidents','Other',
                  gsub('HOME INVASION','Theft',
                  gsub('HUMAN TRAFFICKING','Trafficking',
                  gsub('HUMAN TRAFFICKING - INVOLUNTARY SERVITUDE','Trafficking',
                  gsub('INVESTIGATE PERSON','Other',
                  gsub('Investigate Property','Property',
                  gsub('Larceny','Theft',
                  gsub('Larceny From Motor Vehicle','Theft',
                  gsub('License Violation','Violation',
                  gsub('Liquor Violation', 'Drug Violation',
                  gsub('Medical Assistance','Other',
                  gsub('Missing Person Located','Missing',
                  gsub('Motor Vehicle Accident Response','Vehicle',
                  gsub('Offenses Against Child / Family','Harassment',
                  gsub('Other Burglary','Theft',
                  gsub('Prison Related Incidents','Other',
                  gsub('Property Lost','Property',
                  gsub('Recovered Stolen Property','Property',
                  gsub('Robbery','Theft',
                  gsub('Evading Fare','Theft',
                  gsub('Verbal Disputes','Harassment',
                  gsub('Property Related Damage','Property',
                  gsub('Residential Burglary','Theft',
                  gsub('Search Warrants','Property',
                  gsub('Towed','Vehicle',
                  gsub('Assembly or Gathering Violations','Violation',
                  gsub('Warrant Arrests','Other', bcrime3$Crimes
                  )))))))))))))))))))))))))))))))))))))))))))))))
bcrime3$Crimes = gsub('Ballistics','Weapon',
                 gsub('Burglary - No Property Taken','Theft',
                 gsub('Counterfeiting','Theft',
                 gsub('Drug Violation','Drug Violation',
                 gsub('Explosives','Weapon',
                 gsub('Firearm Violation','Weapon',
                 gsub('Weapons','Weapon',
                 gsub('Weapon Possession','Weapon',
                 gsub('Bomb Hoax','Weapon',
                 gsub('Homicide','Murder',
                 gsub('Investigate Person','Other',
                 gsub('Landlord/Tenant Disputes','Property',
                 gsub('Propertys','Property',
                 gsub('License Plate Related Incidents','Vehicle',
                 gsub('Manslaughter','Murder',
                 gsub('Missing Person Reported','Missing',
                 gsub('Operating Under the Influence','Drug Violation',
                 gsub('Phone Call Complaints','Other',
                 gsub('Property Found','Property',
                 gsub('Prostitution','Trafficking',
                 gsub('Restraining Order Violations','Violation',
                 gsub('Violations','Violation',
                 gsub('Service','Theft',
                 gsub('Vandalism','Property', bcrime3$Crimes))))))))))))))))))))))))  
```
Three frequency tables
```{r}
table(bcrime3$UCR_PART) 

table(bcrime3$Crimes)

table(bcrime3$DAY_OF_WEEK)
```
Two summary measures
```{r}
summary(bcrime3$MONTH)

summary(bcrime3$HOUR)
```
Maximum occurrences
```{r}
bcrime4=bcrime3 %>% filter(!Location=="000000000 000000000")

names(which.max(table(bcrime4$Location)))

names(which.max(table(bcrime4$REPORTING_AREA)))

names(which.max(table(bcrime4$HOUR)))
```
Two-way table of number of UCR_PART crimes in districts
```{r}
bcrime5 <- table(bcrime4$UCR_PART, bcrime4$DISTRICT)
t(bcrime5)
```
Bar graphs of categorical data
```{r}
ggplot(bcrime4, aes(x = Crimes))+geom_bar(fill="violet")+theme(axis.text.x = element_text(angle=90))+labs(title='Most Crimes to Occur')
```
```{r}
ggplot(bcrime4, aes(x = DISTRICT))+geom_bar(fill="maroon")+theme(axis.text.x = element_text(angle=90))+labs(title='Districts With Most Crime')
```
```{r}
ggplot(bcrime4, aes(x = YEAR))+geom_bar(fill="dark green")+theme(axis.text.x = element_text(angle=90))+labs(title='How Many Shootings Normally Occur by Year')
```


Table summary of average month crimes occur grouped by UCR_PART
```{r}
bcrime4%>%group_by(UCR_PART)%>%summarize(Month=mean(MONTH, na.rm=T))
```
Boxplot, facet-wrapped graph, gradient bar graph, and heat map
```{r}
ggplot(bcrime4,aes(x=HAD_SHOOTING, y=YEAR))+geom_boxplot(fill="lightslateblue")+labs(title='When Shootings Normally Occur by Year')
```
```{r}
ggplot(bcrime4,aes(x=Crimes, fill=HAD_SHOOTING))+geom_bar(position="fill")+facet_wrap(~DAY_OF_WEEK)+theme(axis.text.x=element_text(angle=90))+ labs(title='What Day of Week Most Shootings Occur for a Certain Crime')
```
```{r}
ggplot(bcrime4,aes(x=UCR_PART, y=HOUR))+geom_boxplot(fill=brewer.pal(5, "Set1"))+labs(title='What Hour Crimes of Various Severity are Most Commited')
```
```{r} 
ggplot(data=bcrime4,aes(x=reorder(DISTRICT,YEAR),y=YEAR)) + 
  geom_bar(stat ='identity',aes(fill=YEAR))+
  coord_flip() + 
  theme_grey() + 
  scale_fill_gradient(name="YEAR")+
  labs(title = 'Understanding Which Districts Had Most Crime Over Years',
       y='Year',x='District')
```
```{r}
ggplot(bcrime4,aes(x=UCR_PART,y=HAD_SHOOTING, fill=HOUR))+geom_tile(color="white")+scale_fill_distiller(palette = "Spectral",direction = 1)+theme(axis.text.x = element_text(angle=35))
```


Convert HAD_SHOOTING variable to numeric
```{r}
bcrime5 = bcrime4%>%mutate(HAS_SHOOTING=ifelse(HAD_SHOOTING>0,1,0))%>%select(-HAD_SHOOTING)
```
Multiple linear regression model
```{r}
Modeldata1=bcrime5%>%select(HAS_SHOOTING,HOUR,YEAR)
model1=lm(HAS_SHOOTING~.,Modeldata1)
summary(model1)
```





