---
title: "Final Project"
author: "Sarah Raza (sr6bf)"
date: "12/8/2021"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(RColorBrewer)
marketing_campaign= read.delim("~/Downloads/marketing_campaign.csv", header=FALSE)
```

Data Removal Code
```{r}
market = marketing_campaign%>%select(-V1, -V10, -V11, -V12, -V13, -V14, -V15, -V16, -V18, -V19, -V20, -V21, -V22, -V23, -V24, -V25, -V26, -V27, -V28, -V29)
```

Renaming Columns
```{r}
market1=market%>%rename(Year_birth=V2, Education=V3, Marital_Status=V4, Income=V5, Kidhome=V6, TeenHome=V7, Dt_Customer=V8, Recency=V9, NumWebPurchases=V17)
```

Removing 1st row
```{r}
market2 = market1[-1,]
```

Changing columns to numeric
```{r}
market3 = market2%>%mutate(kid=as.numeric(Kidhome),teen=as.numeric(TeenHome))%>%select(-Kidhome, -TeenHome)
```

Combining kids and teens
```{r}
market4 = market3%>%mutate(Have_Kids = kid + teen)%>%select(-kid,-teen)
```

Changing kids to categorical
```{r}
market5=market4%>%mutate(Kids=ifelse(Have_Kids>0, "Yes", "No"))%>%select(-Have_Kids)
```

Changing the date of enrollment to month of enrollment
```{r}
MarketData=market5%>%mutate(Month=ifelse(grepl("-01-", Dt_Customer), "January", ifelse(grepl("-02-", Dt_Customer), "February", ifelse(grepl("-03-", Dt_Customer), "March", ifelse(grepl("-04-", Dt_Customer), "April", ifelse(grepl("-05-", Dt_Customer), "May", ifelse(grepl("-06-", Dt_Customer), "June", ifelse(grepl("-07-", Dt_Customer), "July", ifelse(grepl("-08-", Dt_Customer), "August", ifelse(grepl("-09-", Dt_Customer), "September", ifelse(grepl("-10-", Dt_Customer), "October", ifelse(grepl("-11-", Dt_Customer), "November", "December"))))))))))))%>%select(-Dt_Customer)
```

Revising Marital Status Answers
```{r}
MarketDat=MarketData%>%mutate(MaritalStatus=ifelse(grepl("Divorced", Marital_Status), "Divorced", ifelse(grepl("Married", Marital_Status), "Married", ifelse(grepl("Together", Marital_Status), "Dating", ifelse(grepl("Widow", Marital_Status), "Widow", "Single")))))%>%select(-Marital_Status)
```

Changing character variables to numeric
```{r}
Customer_Analysis= MarketDat%>%mutate(YearBirth=as.numeric(Year_birth), AnnualIncome=as.numeric(Income), Recent=as.numeric(Recency), WebPurchases=as.numeric(NumWebPurchases))%>%select(-Year_birth, -Income, -Recency, -NumWebPurchases)
```

Categorical summary measures
```{r}
table(Customer_Analysis$Education)

table(Customer_Analysis$MaritalStatus)

table(Customer_Analysis$Kids)

table(Customer_Analysis$Month)
```

Numerical summary measures
```{r}
summary(Customer_Analysis$YearBirth)

summary(Customer_Analysis$AnnualIncome)

summary(Customer_Analysis$Recent)

summary(Customer_Analysis$WebPurchases)
```

Summary of Numerical Variables by Groups
```{r}
Customer_Analysis%>%group_by(Kids)%>%summarize(AvgIncome=mean(AnnualIncome, na.rm=T))

Customer_Analysis%>%group_by(MaritalStatus)%>%summarize(AvgRecent=mean(Recent, na.rm=T))
```

Making Months Appear in Order
```{r}
MarketMonth=Customer_Analysis%>%mutate(Months=factor(Month,levels=c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")))
```


Categorical graphs
```{r}
ggplot(MarketMonth, aes(x = Months))+geom_bar(fill="violet")+theme(axis.text.x = element_text(angle=35))

ggplot(Customer_Analysis, aes(x=Education))+geom_bar(fill="light blue")

ggplot(Customer_Analysis, aes(x=Kids))+geom_bar(fill="aquamarine")

ggplot(Customer_Analysis, aes(x=MaritalStatus))+geom_bar(fill="pink")
```

Numerical Graphs
```{r}
ggplot(Customer_Analysis,aes(y=YearBirth))+geom_boxplot(fill="indianred2")+coord_flip()

ggplot(Customer_Analysis,aes(y=Recent))+geom_boxplot(fill="darkseagreen2")+coord_flip()

ggplot(Customer_Analysis,aes(y=AnnualIncome))+geom_boxplot(fill="cornsilk2")+coord_flip()

ggplot(Customer_Analysis,aes(y=WebPurchases))+geom_boxplot(fill="tan3")+coord_flip()
```

Side by Side Boxplots
Education and Annual Income
```{r}
ggplot(Customer_Analysis,aes(x=Education, y=AnnualIncome))+geom_boxplot(fill=brewer.pal(5, "Set1"))
```

Kids and Web Purchases
```{r}
ggplot(Customer_Analysis,aes(x=Kids, y=WebPurchases))+geom_boxplot(fill="lightslateblue")
```

Scatterplots:
Annual Income and Web Purchases
```{r}
ggplot(Customer_Analysis, aes(x=AnnualIncome, y=WebPurchases))+geom_point(color="springgreen2", alpha=0.5)
```

Birth Year and Recent
```{r}
ggplot(Customer_Analysis,aes(x=YearBirth,y=Recent, color=YearBirth))+geom_point(alpha=0.5)
```

3 or More Variable Graphs:
Facet-Wrap: Marital Status, Kids, Eductaion
```{r}
ggplot(Customer_Analysis,aes(x=MaritalStatus, fill=Kids))+geom_bar(position="fill")+facet_wrap(~Education)+theme(axis.text.x=element_text(angle=90))
```

Bubble Plot: Birth Year, Web Purchases, Kids, Annual Income
```{r}
ggplot(Customer_Analysis, aes(x=YearBirth, y=WebPurchases,size=AnnualIncome, color=Kids))+geom_point(alpha=0.5)
```

Special Graph:
HeatMap
```{r}
ggplot(Customer_Analysis,aes(x=MaritalStatus,y=Kids, fill=WebPurchases))+geom_tile(color="white")+scale_fill_distiller(palette = "Spectral",direction = 1)+theme(axis.text.x = element_text(angle=35))
```

One-Sample Inference:
Annual Income Mean of Customers
```{r}
t.test(Customer_Analysis$AnnualIncome,conf.level = 0.95)
```

Proportion of Customers who have Kids
```{r}
prop.test(1602,2240,conf.level=0.95)
```

Two-Sample Inference:
No kids with a PhD or Kids with a PhD
```{r}
prop.test(c(134, 352), c(638,1602), conf.level = 0.95)
```

College Degree and Married or College Degree and Divorced
```{r}
prop.test(c(231, 431), c(232,864), conf.level = 0.95)
```

Multiple Regression Model:
```{r}
Modeldata=Customer_Analysis%>%select(YearBirth,AnnualIncome,Recent,WebPurchases)
model=lm(AnnualIncome~.,Modeldata)
summary(model)
PredictAnnual_Income<-data.frame(YearBirth= 2001,Recent= 30,WebPurchases= 5 )
predict(model,PredictAnnual_Income,interval="predict")
```

Logistic Regression Model:
```{r}
LogitData=Customer_Analysis%>%select(YearBirth, Recent, AnnualIncome,WebPurchases,Kids)%>%mutate(kid=ifelse(grepl("Yes", Kids), 1, 0))%>%select(-Kids)
logit.model=glm(kid~.,data=LogitData,family="binomial")
summary(logit.model)
Prediction=data.frame(YearBirth=2002,Recent=43,AnnualIncome=30000,WebPurchases=3)
predict(logit.model,Prediction,type="response")
```











